### 2018-02-03 thesis/contents/tn_sagit.tex

\chapter{Study IV: Gaussian Process Classification of Trigeminal Neuralgia using Merged Tractography}
\chaptermark{Study IV}
\label{section:study4}

\section{Abstract}
\textbf{Introduction:} Trigeminal Neuralgia (TN) is a debilitating facial neuropathic pain syndrome involving the trigeminal sensory pathway. We demonstrate the use of a fully-automated group white-matter tractography and Gaussian Process (GP) classification to pinpoint key white-matter diffusivity changes that maximally differentiate patients and healthy controls.

\textbf{Methods:} We used SAGIT based group merged tractography for 36 sex-matched TN patients (right-sided pain) and 36 controls, examining the following white matter pathways: trigeminal nerve (CN V), pontine decussation, and thalamocortical fibers. GP classifiers were trained by scrolling a moving window over CN V and S1 tractography centroids. FA, GFA, RD, AD, and MD metrics were evaluated for both groups, analysing TN vs. Control and affected vs. unaffected groups. Classifiers that performed at greater-or-equal-to 75\% accuracy were accepted.

>>>>

\chapter{Study IV: Gaussian Process Classification of Trigeminal Neuralgia using Merged Group Tractography}
\chaptermark{Study IV}
\label{section:study4}

\section{Abstract}
\textbf{Introduction:}  Imaging of trigmeinal neuralgia (TN) has demonstrated key diffusivity alterations in the trigeminal nerve but the difficulties of group tractography analysis have previously prevented us from assessing the entire trigeminal pathway. We apply fully-automated group white-matter tractography to now image the trigeminal sensory pathway, and Gaussian Process (GP) classification to pinpoint key white-matter diffusivity changes that maximally differentiate TN in patients and healthy controls.

\textbf{Methods:} We used SAGIT based group merged tractography for 36 sex-matched TN patients (right-sided pain) and 36 controls, examining the following white matter pathways: trigeminal nerve (CN V), pontine decussation, and thalamocortical fibers. GP classifiers were trained by scrolling a moving window over CN V and S1 tractography centroids. FA, GFA, RD, AD, and MD metrics were evaluated for both groups, analysing TN vs. Control and affected vs. unaffected groups. Classifiers that performed at greater-or-equal-to 70\% accuracy were accepted.
